An Automatic Verification Framework for Software Configurations

### 2016-05-10 OSDI/learner.tex
update learner

As a result of this restriction, 
these efforts only consider a rule if it holds over exactly every file in 
the learning set. This behavior can be formally described as follows:
\begin{small}
\begin{align*}
C :=&\ \text{Correct Learning Set}\\
LR :=&\ \text{Learned Rules} :: \{\textrm{Rule}\}\\
RR :=&\ \text{Reported Rules} :: \{\textrm{Rule}\}\\ 
C =&\ \text{\{Configuration Files in Intermediate Representation\}}\\
LR =&\ \{ r\ \mid \forall file \in C,\ holds(r,file)\} \\
RR =&\ \{ r\ \mid r \in LR \ \land \neg\ holds(r,\textrm{user file}) \}\\
\end{align*}
\end{small}
In the above, the ":: \{Rule\}" notation indicates that $LR$, the learned rule set, 
and $RR$, the reported rule set, are sets of rules.
A rule is only committed to the $LR$ learned rule set if 
the rule holds on all files in the training set $C$, as denoted by the $\forall$ quantifier. 
A broken rule is reported to the user in $RR$ rule set if the rule is in $LR$, 
but does not hold on the input user file.

Each rule can be viewed as a mapping from 
entry keys $(k_1, \ldots, k_n)$ in the intermediate representation of a 
configuration file to a boolean function, $B_{rel}$. This can be represented by a rule
$r := (k_1, \ldots, k_n) \rightarrow B_{rel}$.
$B_{rel}$ is a Boolean function specific to the relation to assess, meaning that 
the desired relation can either be satisfied or not satisfied. 
For instance, for the rule that {\tt max\_connections} must 
be greater than {\tt mysql.max\_persistent}, the rule is denoted by
({\tt max\_connections}, {\tt mysql.max\_persistent}) $\to 1$, where 
$B_{>}$ is set to 1.

@@

    \State $R = \{\}$ // rule set
    \State $P$ // probabilities associated with rule set
    \For{each file $F$ in $C$}
      \For{each $\tau \in \theta$}
        \State $Q = filter(F,\tau)$ 
        \State // $Q$ contains entries in $F$ with $\tau$ in type list
        \State $RM$ contains entries in remove set
        \For{each entry $e$ in $Q$}
               \If{$g(e)$ is false}
                        \State continue to next entry
                        \Else \enspace $CR = \textrm{constructRules}(e, \tau, Q \backslash RM)$
                        \State $R = R  \cup CR$
                        \State For each $r \in CR$, update $P$ 
               \EndIf
               \State $RM = RM \cup e$
        \EndFor
     \EndFor
    \EndFor
    \State Based on $P$, accept rules in $R$ according to threshold. 
  \end{algorithmic}
\end{algorithm}

@@

Our probabilistic approaches for learning the entry missing, 
ordering, and fine-grained value correlation rules stem 
from existing work with building 
the non-probabilistic versions of these rule-learning algorithms. 
We start with the idea that for each of these rules, 
we are going to consider all possible pairs of keys that appear in every 
file, and for our learning process, calculate the likelihood that each of 
these pairs constitute a rule. 
In essence, this is a mechanism that will consider many different possible 
pairs of lines of entries in the file, 
and attempt to compile a set of these pairs 
that are expressed more often than others as a basis for finding patterns 
within the example set that can be used to evaluate new files.

More formally, this approach can be defined as follows. 
The output of our learning algorithm is augmented to a map from 
key-value pairs to a probability distribution over number of possible 
relations.

\[
\{ P\_Rule = (a_j, a_k) | j \neq k \} \rightarrow \{ (R_1, R_2, ... , R_n) \}
\]

We can think of each of these $(a_j, a_k)$ as possibly having a different relationship, defined by the set $\{ R_i \}$, which cover the entire outcome space of possible relationships between the two values $(a_j, a_k)$.

For the entry missing rules, we define $R_1$ as the event that $a_j$ and
$a_k$ appear together, and $R_2$ to be the event that $a_j$ appears
without $a_k$, or by the transitive equivalent, $a_k$ appears without
$a_j$. For the ordering rules, we define $R_1$ as the event that
$a_j$ appears before $a_k$ and $R_2$ be the event that $a_k$ appears
before $a_j$. For the value correlation rules, we define $R_1$ as the
event that $a_j \leq a_k$, $R_2$ the event that $a_j = a_k$, and $R_3$
the case that $a_j \geq a_k$. Notice that the $R_i$ do not have to be
disjoint, but only have to union to the entire probability space.

By examining the learning set, we will derive a distribution of the set $\{R_i\}$ based on how many times we observe an occurrence of each relation. This distribution will then be used at checking time to determine if a user's configuration has broken a likely rule.

\begin{small}
\begin{flalign*}
I\ \ =&\ \text{Incorrect Learning Set}\\
\text{::}\ & \text{\{Configuration Files in Intermediate Representation\}}\\
LP =&\ \text{Learned Probabilistic Rules :: \{P\_Rule\}}\\
RP =&\ \text{Reported Probabilistic Rules :: \{P\_Rule\}}\\
LP =&\ \text{count\_relation\_occurrences}(I)\\
RP =&\ \{ r\ \mid r \in LP\ \land \Pi(r)>p \land \neg r(userfile) \}
\end{flalign*}
\end{small}

A rule will be reported as broken if the probability the rule is
correct, $\Pi$, is greater than some user defined constant, $p$. This
constant can be adjust to the user's preference. A small $p$ will
increase the likelihood of finding an error, but also increase the
number of false positives that are reported.

>>>>

As a result of this restriction, 
these efforts only consider a rule if it holds over exactly every file in 
the learning set. This behavior can be formally described as follows:
\begin{small}
\begin{align*}
C :=&\ \text{Correct Learning Set}\\
LR :=&\ \text{Learned Rules} :: \{\textrm{Rule}\}\\
RR :=&\ \text{Reported Rules} :: \{\textrm{Rule}\}\\ 
C =&\ \text{\{Configuration Files in Intermediate Representation\}}\\
LR =&\ \{ r\ \mid \forall file \in C,\ holds(r,file)\} \\
RR =&\ \{ r\ \mid r \in LR \ \land \neg\ holds(r,\textrm{user file}) \}\\
\end{align*}
\end{small}
In the above, the ":: \{Rule\}" notation indicates that $LR$, the learned rule set, 
and $RR$, the reported rule set, are sets of rules.
A rule is only committed to the $LR$ learned rule set if 
the rule holds on all files in the training set $C$, as denoted by the $\forall$ quantifier. 
A broken rule is reported to the user in the $RR$ rule set if the rule is in $LR$, 
but does not hold on the input user file.

In general, each rule can be viewed as a mapping from 
entry keys $(k_1, \ldots, k_n)$ in the intermediate representation of a 
configuration file to a probability function, $P_{rel}$. This can be represented by a rule
$r := (k_1, \ldots, k_n) \rightarrow P_{rel}$.
$P_{rel}$ is a probability function specific to the relation to assess.
The problem with the strawman solution is that the probability functions are restricted to be 
boolean, which is to say that the relation can either be unsatisfied or satisfied, either 0 or 1. 
For instance, for the rule that {\tt max\_connections} must 
be greater than {\tt mysql.max\_persistent}, the rule is denoted by
({\tt max\_connections}, {\tt mysql.max\_persistent}) $\to 1$, where 
$B_{>}$ is set to 1. In our approach, we permit the probability functions to take on any value from 0 to 1.

@@

    \State $R = \{\}$ // rule set
    \State $\Pi$ // probabilities associated with rule set
    \For{each file $F$ in $C$}
      \For{each $\tau \in \theta$}
        \State $Q = filter(F,\tau)$ 
        \State // $Q$ contains entries in $F$ with $\tau$ in type list
        \State $RM$ contains entries in remove set
        \For{each entry $e$ in $Q$}
               \If{$g(e)$ is false}
                        \State continue to next entry
                        \Else \enspace $CR = \textrm{constructRules}(e, \tau, Q \backslash RM)$
                        \State $R = R  \cup CR$
                        \State For each $r \in CR$, update $\Pi$ 
               \EndIf
               \State $RM = RM \cup e$
        \EndFor
     \EndFor
    \EndFor
    \State return $(\Pi,R)$
  \end{algorithmic}
\end{algorithm}

@@

Our probabilistic approaches for learning missing entry, 
ordering, and fine-grained value correlation rules stem 
from existing work with building 
non-probabilistic versions of these rule-learning algorithms. 
For each of these rules, 
we consider all possible permutations of keys that appear in every 
file which are appropriately typed, and for our learning process, calculate the likelihood that each of 
these permutations constitute a rule. 
Rule patterns in the example set that appear frequently are accepted,
so they can be used to evaluate new files.

More formally, this approach can be defined as follows:

\begin{small}
\begin{flalign*}
I\ \ :=&\ \text{Incorrect Learning Set}\\
LP :=&\ \text{Learned Probabilistic Rules :: \{P\_Rule\}}\\
RP :=&\ \text{Reported Probabilistic Rules :: \{P\_Rule\}}\\
I =&\  \text{\{Configuration Files in Intermediate Representation\}}\\
(\Pi, LP) =&\ \text{learnRules}(I)\\
RP =&\ \{ r\ \mid r \in LP\ \land \Pi(r)>p \land \neg holds(r, userfile) \}
\end{flalign*}
\end{small}

%this approach can be defined in Algorithm~\ref{alg:crules}.

%More formally, this approach can be defined in Algorithm~\ref{alg:plearn}. 
Note that the probability set associated with the rules, $\Pi$, contains values
ranging from 0 to 1, which are then compared against an acceptance threshold for the final output.
%\[
%\{ P\_Rule = (a_j, a_k) | j \neq k \} \rightarrow \{ (R_1, R_2, ... , R_n) \}
%\]

%We can think of each of these $(a_j, a_k)$ as possibly having a different relationship, defined by the set $\{ R_i \}$, which cover the entire outcome space of possible relationships between the two values $(a_j, a_k)$.

%For the entry missing rules, we define $R_1$ as the event that $a_j$ and
%$a_k$ appear together, and $R_2$ to be the event that $a_j$ appears
%without $a_k$, or by the transitive equivalent, $a_k$ appears without
%$a_j$. For the ordering rules, we define $R_1$ as the event that
%$a_j$ appears before $a_k$ and $R_2$ be the event that $a_k$ appears
%before $a_j$. For the value correlation rules, we define $R_1$ as the
%event that $a_j \leq a_k$, $R_2$ the event that $a_j = a_k$, and $R_3$
%the case that $a_j \geq a_k$. Notice that the $R_i$ do not have to be
%disjoint, but only have to union to the entire probability space.

%By examining the learning set, we will derive a distribution of the set $\{R_i\}$ based on how many times we observe an occurrence of each relation. This distribution will then be used at checking time to determine if a user's configuration has broken a likely rule.

A rule $r$ will be reported as broken if the probability the rule is
correct, $\Pi(r)$, is greater than some user defined constant, $p$. This
constant can be adjust to the user's preference. A small $p$ will
increase the likelihood of finding an error, but also increase the
number of false positives that are reported.
